---
title: "ViscousFlow"
author: "Christina Bogner"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: github_document
bibliography: [inst/REFERENCES.bib]
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## Package aims

<!-- badges: start -->
<!-- badges: end -->

This package analyses laboratory irrigation experiments on soil columns. It uses the viscous flow approach by @Germann2018.

## Installation

You can install ViscousFlow from GitHub with `devoolts' like so:

``` r
# install.packages("devtools")
devtools::install_github("ChrisBogner/ViscousFlow")
```
## Example

The package contains two data sets, namely `tracer` and `drainage`. Both originate from the publication by @Bogner2019 (there called column C1, irrigation intensity 10 mm h$^{-1}$.)

This is a basic example which shows you how to solve a common problem: TODO

```{r example}
library(ViscousFlow)
## basic example code
```

## Refereneces
<div id="refs"></div>
