---
title: "ViscousFlow"
author: "Christina Bogner"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
    github_document:
      #pandoc_args: --webtex=https://ibm.codecogs.com/png.latex?
#always_allow_html: yes
bibliography: [inst/REFERENCES.bib]
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## Package aims

<!-- badges: start -->
<!-- badges: end -->

This package analyses laboratory irrigation experiments on soil columns. It uses the viscous flow approach by @Germann2018.

## Installation

You can install ViscousFlow from GitHub with `devoolts' like so:

``` r
# install.packages("devtools")
devtools::install_github("ChrisBogner/ViscousFlow")
```
## Example

This is a basic example which shows you how to fit the viscous flow equation to the decreasing limb of a drainage curve and to calculate the viscous flow parameters $F$ and $L$. 

The package contains two data sets, namely `tracer` and `drainage`. Both originate from the publication by @Bogner2019. There, it is called column C1 and was packed from loose soil material collected form a forest soil in southeast Germany (50°08'32.8'' N 11°51'52.9'' E). Prior to the irrigation experiment, the soil column was saturated from below and then drained to field capacity. It was irrigated at 10 mm h^-1^ during 64410 sec (17.9 h). During the experiment, a suction of -10 hPa was applied at the bottom of the soil column to prevent saturation. The data set `drainage' contains the drainage from this soil column C1. The data was smoothed (see the original open access publication @Bogner2019 for details).

The irrigation water contained Bromide that was measured in the drainage water with an ion-sensitive electrode. The data set `tracer' contains the normalized Bromide concentration (i.e. concentration in the drainage divided by the concentration in the irrigation water).

Load libraries and data.
```{r example}
library(ViscousFlow)
library(tidyverse)
library(gridExtra)

data(drainage)
data(tracer)
```

Define T~B and T~E, the start and end times of the irrigation experiment.

```{r}
TB <- 0
TE <- 64410
```


Plot the data.
```{r}
g_all <- list(geom_line(),
  xlab('Time since start of irrigataion(sec)'),
  ylab(expression(paste('Drainage (mm ', h^{-1}, ')'))),
  geom_vline(aes(xintercept = TE, col = 'TE'), lty = 2),
  scale_colour_manual(name = ' ', values = ('end of irrigation' = 'blue'),
                      labels = 'end of irrigation'),
  theme(legend.position=c(0.2, 0.9), 
        legend.background = element_rect(fill="transparent", colour=NA),          legend.key = element_rect(fill = "white"))
)

g1 <- ggplot(data = drainage, aes(x = time_sec, y = q_mmh)) + 
  g_all +
  ggtitle('Drainage column C1')

g2 <- ggplot(data = tracer, aes(x = time, y = value)) + 
  g_all +
  ggtitle('Tracer breakthrough column C1')

grid.arrange(g1, g2, ncol = 1)
```


Calculate the breakthrough time of the tracer. The variable `time_interval` reduces the searched time. The breakthroug time is defined as the largest curvature in `time_interval`.
```{r}
find_tracer_breakthrough(tracer_data = tracer, time_interval = c(30000, 40000), do_plot = T)

```


## Refereneces
<div id="refs"></div>
